Industrial, large-scale model predictive control with structured neural networks

### 2021-02-04 paper.tex
more edits.

\subsection{Training}
The following mean squared loss is minimized to determine the weights for the NN
controller
\begin{align}
    J (\theta) = \dfrac{1}{2}\sum_{j=1}^{N_{tr}} 
[\kappa_{NN}(x_j, x_{sj}, u_{sj}; \theta) - \kappa_N(x_j, x_{sj}, u_{sj})]^2 
\end{align}
in which $\theta$ is the set of parameters to be optimized, ($W_i, b_i,
W_{h+1}$) for $i \in \bbI_{1:h}$, subscript $j$ is used to denote the training
sample index, and $N_{tr}$ is the number of training samples. The stochastic
optimization algorithm Adam \citep*{kingma:ba:2014} is used for all the NN
training performed in the case studies in this article. We do not consider any
regularization penalty in the training objective. The recent empirical and
theoretical works in the machine learning literature have shown that NNs with
large number of parameters have good interpolation capabilities
\citep*{belkin:hsu:ma:mandal:2019, zhang:bengio:hardt:recht:vinyals:2017,
arora:simon:hu:li:wang:2019, allen-zhu:li:liang:2019} without any form of
explicit regularization during training. The generalization abilities of NNs for
the architectures considered in this article is examined in the simulations
presented in the Section \ref{sec:simulation_studies}.

@@

The related work by \cite{hertneck:kohler:trimpe:allgower:2018} examines the robustness of nonlinear systems in feedback with approximate MPC controllers. The MPC control law approximation error is treated as an input disturbance, and a robust MPC formulation based on constraint tightening procedures is used to establish the allowable approximation error. The robustness of NN controllers trained with samples of the nominal MPC control law in presence of process disturbances and state estimation errors is not established in that paper.

@@

The validation simulations are performed directly with the plant model for the case studies in this section. For industrial deployment, the NN training should be followed by an offline validation step of the trained network to verify the quality of the NN controller before online deployment. This validation can be performed by: a) quantifying the mean squared error (MSE) of the trained NN controller on test data generated using the optimal MPC, b) examining closed-loop performance metrics in simulations carried out with the linear model used in MPC and the optimal MPC and NN controllers. Additional probabilistic validation techniques for NNs trained using robust MPC solutions for nonlinear problems are proposed in \cite{karg:alamo:lucia:2019} and \cite{hertneck:kohler:trimpe:allgower:2018}.

@@

First, the amount of data required to obtain an accurate NN controller is
studied. To this aim, NNs with the three chosen architectures are trained with
incremental increases of $10^4$ training samples starting from $4 \times 10^4$
to $1.5 \times 10^5$ samples. A batch size of 1024 samples is used in Adam, and
all the NNs are trained for 2000 epochs. The NNs after training are used in
validation simulations with the setpoint and disturbance signals generated for
the expected plant behavior case. The closed-loop performance is compared with
the optimal MPC controller and the $\%$ performance loss obtained is plotted in
Figure \ref{fig:cstrs_statistics} (top left) as a function of the number of
training samples. All the trained NNs provide less than $1 \%$ loss in this
performance metric after training with about $9 \times 10^4$ samples. The best
$\%$ performance loss obtained by the NNs is summarized in Table
\ref{table:cstrs}.

@@

Figure \ref{fig:cdu_statistics} shows the statistics for the simulation study
performed for this example. The data requirement to obtain an accurate NN
controller is shown in the top left. The batch size in Adam was chosen as 2048
samples, and all the NNs are trained for 1500 epochs. The transient controller
performance index for one validation simulation obtained with the three
structured NNs, satK, and MPC controllers is shown in the center bottom. The top
right plot shows the histogram of online computation times of the NNs and QP
solver. Table \ref{table:cdu} gives a summary of the maximum time required to
train the NNs, number of parameters in each NN, memory footprint of the NNs, and
the $\%$ performance loss metric obtained by the structured NNs, SS, satK, and
SH controllers in the validation simulations.

@@

>>>>

\subsection{Training}

The following mean squared error (MSE) is minimized to determine the weights for
the NN controller
\begin{align}
    J (\theta) = \dfrac{1}{2}\sum_{j=1}^{N_{tr}} 
[\kappa_{NN}(x_j, x_{sj}, u_{sj}; \theta) - \kappa_N(x_j, x_{sj}, u_{sj})]^2 
\end{align}
in which $\theta$ is the set of parameters to be optimized, ($W_i, b_i,
W_{h+1}$) for $i \in \bbI_{1:h}$, subscript $j$ is used to denote the training
sample index, and $N_{tr}$ is the number of training samples. The stochastic
optimization algorithm Adam \citep*{kingma:ba:2014} is used for all the NN
training performed in the case studies in this article. We do not consider any
regularization penalty in the training objective. The recent empirical and
theoretical works in the machine learning literature have shown that NNs with
large number of parameters have good interpolation capabilities
\citep*{belkin:hsu:ma:mandal:2019, zhang:bengio:hardt:recht:vinyals:2017,
arora:simon:hu:li:wang:2019, allen-zhu:li:liang:2019} without any form of
explicit regularization during training. The generalization abilities of NNs for
the architectures considered in this article is examined in the simulations
presented in the Section \ref{sec:simulation_studies}.

@@

The related work by \cite{hertneck:kohler:trimpe:allgower:2018} examines the
robustness of nonlinear systems in feedback with approximate MPC controllers.
The MPC control law approximation error is treated as an input disturbance, and
a robust MPC formulation based on constraint tightening procedures is used to
establish the allowable approximation error. The robustness of NN controllers
trained with samples of the nominal MPC control law in presence of process
disturbances and state estimation errors has not been established to date in the
literature.

@@

The validation simulations are performed directly with the plant model for the
case studies in this section. For industrial deployment, the NN training should
be followed by an offline validation step of the trained network to verify the
quality of the NN controller before online deployment. This validation can be
performed by: a) quantifying the MSE of the trained NN controller on test data
generated using the optimal MPC, b) examining closed-loop performance metrics in
simulations carried out with the linear model used in MPC and the optimal MPC
and NN controllers. Additional probabilistic validation techniques for NNs
trained using robust MPC solutions for nonlinear problems are proposed in
\cite{karg:alamo:lucia:2019} and \cite{hertneck:kohler:trimpe:allgower:2018}.

@@

First, the amount of data required to obtain an accurate NN controller is
studied. To this aim, NNs with the three chosen architectures are trained with
incremental increases of $10^4$ training samples starting from $4 \times 10^4$
to $1.5 \times 10^5$ samples. A batch size of 1024 samples is used in Adam, and
all the NNs are trained for 2000 epochs. For all the NN training performed,
10$\%$ of the training dataset is kept as a buffer dataset, which is monitored
during training and the network weights are updated only when the MSE decreases
on this buffer dataset after every training epoch. The NNs after
training are used in validation simulations with the setpoint and disturbance
signals generated for the expected plant behavior case. The closed-loop
performance is compared with the optimal MPC controller and the $\%$ performance
loss obtained is plotted in Figure \ref{fig:cstrs_statistics} (top left) as a
function of the number of training samples. All the trained NNs provide less
than $1 \%$ loss in this performance metric after training with about $9 \times
10^4$ samples. The best $\%$ performance loss obtained by the NNs is summarized
in Table \ref{table:cstrs}.

@@

Figure \ref{fig:cdu_statistics} shows the statistics for the simulation study
performed for this example. The data requirement to obtain an accurate NN
controller is shown in the top left. The batch size in Adam was chosen as 2048
samples, and all the NNs are trained for 1500 epochs. For all the NN training,
the size of the buffer dataset was chosen as 5$\%$ of the training dataset. The
transient controller performance index for one validation simulation obtained
with the three structured NNs, satK, and MPC controllers is shown in the center
bottom. The top right plot shows the histogram of online computation times of
the NNs and QP solver. Table \ref{table:cdu} gives a summary of the maximum time
required to train the NNs, number of parameters in each NN, memory footprint of
the NNs, and the $\%$ performance loss metric obtained by the structured NNs,
SS, satK, and SH controllers in the validation simulations.

@@

Future research in the area of approximating MPC control laws using NNs can
focus on developing methods for approximating the target selector QP solution,
systematically handling state constraints, extensions to nonlinear MPC problems,
and establishing conditions for guaranteed closed-loop stability with NN
controllers.
